---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# DEBBI

<!-- badges: start -->
<!-- badges: end -->

The primary goal of DEBBI (short for 'Differential Evolution-based Bayesian Inference') is to provide efficient access to and enable reproducible use of Bayesian inference algorithms such as Differential Evolution Markov Chain Monte Carlo, Differential Evolution Variation Inference, and Differential Evolution maximum a posteriori estimation. The second goal of this package is to be compatible with likelihood-free Bayesian inference methodologies that use approximations of the likelihood function such as probability density approximation, kernel-based approximate Bayesian computation, and synthetic likelihood.

## Installation

You can install the development version of DEBBI from [GitHub](https://github.com/bmgaldo/DEBBI) with:

``` r
# install.packages("devtools")
devtools::install_github("bmgaldo/DEBBI")
```

## Example

Estimate mean parameters of two independent normal distributions with known standard deviations using DEMCMC

```{r example}
library(DEBBI)


# simulate from model
dataExample=matrix(rnorm(100,c(-1,1),c(1,1)),nrow=50,ncol=2,byrow = T)

# list parameter names
par_names_example=c("mu_1","mu_2")

# log posterior likelihood function = log likelihood + log prior | returns a scalar
LogPostLikeExample=function(x,data,par_names){
  out=0
  
  names(x)<-par_names
  
  # log prior
  out=out+sum(dnorm(x["mu_1"],0,sd=1,log=T))
  out=out+sum(dnorm(x["mu_2"],0,sd=1,log=T))
  
  # log likelihoods
  out=out+sum(dnorm(data[,1],x["mu_1"],sd=1,log=T))
  out=out+sum(dnorm(data[,2],x["mu_2"],sd=1,log=T))
  
  return(out)
}

# Sample from posterior
DEMCMC(LogPostLike=LogPostLikeExample,
                  control_pars=AlgoParsDEMCMC(n_pars=length(par_names_example),
                                     n_iter=1000, n_chains=12,
                                     init_sd=.01,init_center=0,
                                     n_cores_use=1,step_size=NULL,
                                     jitter_size=1e-6,parallelType = "none"),data=dataExample,
                  par_names = par_names_example)

par(mfrow=c(1,3))
for(i in 1:length(par_names_example)){
  
matplot(out$samples[,,i],type='l',main=par_names_example[i])
  
}

matplot(out$log_post_density,type='l')


```
